Update and evaluation of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-12-06 DryDep_supplement.tex
Suggestions by Frode - Part I.
Rearangement of supplement. Minor changes in the text. Almost finished proofreading.

\documentclass[manuscript]{copernicus}
\graphicspath{{pictures/}}          % graphics
\begin{document}
\clearpage
\setcounter{page}{1}

\section*{Supplement}
\subsection*{S.1 Dynamic viscosity of air and quasi-laminar resistance}
\subsubsection*{S.1.1 Sutherland's law and linear fits}
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity.pdf}
\end{center}
\subsubsection*{S.1.2 Comparison of $z_o$ and $R_b$ for $\mu(T) = m\cdot T$ and Sutherland's law: \chem{H_2O}}
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity_zo_Rb.pdf}
\end{center}
\subsubsection*{S.1.3 $R_b$ for different species}
\appendixfigures
\begin{center}
  \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_1.pdf}
  \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_2.pdf}
\end{center}

%\begin{minipage}[t]{1.\textwidth}
\subsection*{S.2 Tabulated parameters for stomatal conductance}
\appendixtables
\begin{sidewaystable*}[!htbp]
  %\caption{S.2 Tabulated parameters for stomatal conductance}

@@

@@

    Code$^*$ & $D_\text{max}$ & $D_\text{min}$ & $D_\text{crit}$ & $R_\chem{SO}$ & $R_\chem{O_3}$ & $h$ & $d_\text{SGS}$ & $d_\text{EGS}$ & $\nabla d_\text{SGS}$ & $\nabla d_\text{EGS}$ & $\text{LAI}_\text{min}$ & $\text{LAI}_\text{max}$ & LS & LE \\
    &(\unit{kPA})&(\unit{kPa})&(\unit{kPa})&(\unit{s\,m^{-1}})&(\unit{s\,m^{-1}})&(\unit{m})&(\unit{day})&(\unit{day})&(\unit{days/^\circ lat})&(\unit{days/^\circ lat})&(\unit{m^2\,m^{-2}})&(\unit{m^2\,m^{-2}})&(\unit{days})&(\unit{days}) \\

@@

    CF & 0.5 & 3 & 1000 & 0 & 200 & 20 & 0 & 366 & 0 & 0 & 5 & 5 & 1 & 1\\
    DF & 1 & 3.25 & 1000 & 0 & 200 & 20 & 100 & 307 & 1.5 & -2.0 & 0 & 4 & 20 & 30\\
    NF & 1 & 3.2 & 1000 & 0 & 200 & 8 & 0 & 366 & 0 & 0 & 4 & 4 & 1 & 1\\
    BF & 2.2 & 4 & 1000 & 0 & 200 & 15 & 0 & 366 & 0 & 0 & 4 & 4 & 1 & 1\\
    TC & 1.2 & 3.2 & 8 & 0 & 200 & 1 & 123 & 213 & 2.57 & 2.57 & 0 & 3.5 & 70 & 20\\
    MC & 1 & 2.5 & 1000 & 0 & 200 & 2 & 123 & 237 & 2.57 & 2.57 & 0 & 3 & 70 & 44\\
    RC & 0.31 & 2.7 & 10 & 0 & 200 & 1 & 130 & 250 & 0 & 0 & 0 & 4.2 & 35 & 65\\
    SNL & 1.3 & 3 & 1000 & 0 & 400 & 0.5 & 0 & 366 & 0 & 0 & 2 & 3 & 192 & 96\\
    GR & 1.3 & 3 & 1000 & 0 & 1000 & 0.3 & 0 & 366 & 0 & 0 & 2 & 3.5 & 140 & 135\\
    MS & 1.3 & 3.2 & 1000 & 0 & 200 & 2 & 0 & 366 & 0 & 0 & 2.5 & 2.5 & 1 & 1\\
    WE  & 1 & 0 & 1000 & 50 & 400 & 0.5 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
    TU & 1 & 0 & 1000 & 500 & 400 & 0.5 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
    DE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
    W  & 1 & 0 & 1000 & 1 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
    ICE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\
    U & 1 & 0 & 1000 & 400 & 400 & 10 & 0 & 366 & 0 & 0 & 0 & 0 & 0 & 0\\

@@

\end{sidewaystable*}
%\end{minipage}

\subsection*{S.3 Latitude dependent vegetation height}
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight]{pictures/tree_height.pdf}
\end{center}

\subsection*{S.4 Temperature dependent greening season}
\subsubsection*{S.4.1 Python code snippets}
\begin{verbatim}
#----------------------------------------------------------------------------------
import numpy as np
import pandas as pd
import xarray as xr
#----------------------------------------------------------------------------------
def start_growing_season_fixed(lat, **kwargs):
    '''
    Begin of growing season parameterization adapted from SMOKE-BEIS.
    '''
    bLeap = kwargs.pop('leap', False)
    if (bLeap):
       MAY31 = 152
       NOV1  = 306
       DEC31 = 366
    else:
       MAY31 = 151
       NOV1  = 305
       DEC31 = 365

@@

    return({'sgs':start_gs,'egs':end_gs})
#----------------------------------------------------------------------------------
def growing_season_stadyn(xr_temp):
    '''
    Preprocess the greening season for OsloCTM3.
    Setting using the 5deg-5days criteria and in case this fails
    or is out of the defined bounds,
    fall back to SMOKE-BEIS parameterization.
    '''
    # Fetch leap year from input data
    bLeap = False
    if xr_temp.time.size==366:
       bLeap = True
    # Use 5deg-5days criteria   
    if (xr_temp.lat > 45 and xr_temp.lat < 85):
        gs = growing_season(xr_temp)
        sgs = (gs['sgs'][0],)
        egs = (gs['egs'][0],)
        if ( not gs['sgs'][1] or not gs['egs'][1] or sgs[0]>=egs[0]):
            # Check for failing 5deg-5days criteria => Fall back to fixed
            sgs = start_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
            egs = end_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
    elif (xr_temp.lat < -35 and xr_temp.lat >= -65):
        gs = growing_season(xr_temp, sh=True)
        sgs = (gs['sgs'][0],1)
        egs = (xr_temp.size, gs['egs'][0])
        # Check for failing 5deg-5days criteria => Fall back to fixed
        if (not gs['sgs'][1] or not gs['egs'][1] or sgs[0]>=egs[0]):
            sgs = start_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
            egs = end_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
    # Use SMOKE-BEIS parameterization
    else:
        sgs = start_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
        egs = end_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
    # Generate GDAY and GLEN fields
    # in case no growing season has been allocated
    if (sgs[0]==egs[0]==0):
        gday = np.repeat(0,len(xr_temp.time))
        glen = 0
    # in normal cases
    else:
        # handle northern hemisphere
        if len(sgs)==1:
            gday = np.concatenate((np.repeat(0,int(sgs[0])-1),
                                   np.arange(1,int(egs[0])-int(sgs[0])+1),
                                   np.repeat(0,len(xr_temp.time)-
                                   int(egs[0])+1)))
            glen = int(egs[0])-int(sgs[0])+1
        # handle southern hemisphere
        else:
            gday = np.concatenate((np.arange(1,int(egs[1])+1)+len(xr_temp.time)
                                   -int(sgs[0])+1,
                                   np.repeat(0,int(sgs[0])-int(egs[1])),
                                   np.arange(1,len(xr_temp.time)-int(sgs[0])+1)))
            glen = len(xr_temp.time)-(int(sgs[0])-int(egs[1]))+1
    # This should not happen, but in case it does, print info.
    if not(len(gday)==len(xr_temp.time)):
        print(xr_temp, sgs, egs)
    # Output
    data_gday = xr.DataArray(gday.astype(int),[('time', xr_temp.time.data)])
    return(data_gday, (int)(glen))
#----------------------------------------------------------------------------------
\end{verbatim}
%\subsubsection*{S.4.2 GLEN: NH mid--high latitudes only}
%\appendixfigures
%\begin{center}
%  \includegraphics[height=0.4\textheight, clip, trim={0.cm 1.75cm 0.cm 1.75cm}]{pictures/GLEN_2005.pdf}
%\end{center}

\subsection*{S.5 De-accumulation of photosynthetic active radiation (PAR) from OpenIFS: January 2nd 2005}
\subsubsection*{S.5.1 Output from OpenIFS: Accumulated PAR}
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.65cm}]{pictures/PAR_as_is.png}
\end{center}
\subsubsection*{S.5.2 Partly de-accumulated}
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.65cm}]{pictures/PAR_partly_deaccumulated.png}
\end{center}
\subsubsection*{S.5.3 De-accumulation of hour 00}
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.65cm}]{pictures/PAR_deaccumulating_00h.png}
\end{center}

\subsection*{S.6 Normed zonal average ozone dry deposition velocity}
%\subsubsection*{S.6.1 }
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-norm_total_ozone_drydepvelo}
\end{center}

>>>>

\documentclass[manuscript]{copernicus}
\usepackage{graphicx}
\usepackage{subcaption}
%\usepackage{subfigure}
\graphicspath{{pictures/}}          % graphics
\begin{document}
\clearpage
\setcounter{page}{1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplement}
\subsection*{S.1 Aerodynamical resistance}
\begin{table}[!htbp]
  \caption{Molecular diffusivity for a gas $i$ given as ration $D_\chem{H_2O}/D_i$, reactivity $f^i_0$, and effective Henry's Law constant $H^i_*$.}
  \begin{tabular}{lcrr@{$\cdot$}l}
    \tophline
    & $D_\chem{H_2O}/D_i$ & $f^i_0$ & \multicolumn{2}{l}{$H^i_*$}\\
    & & & \multicolumn{2}{l}{(\unit{M\,atm^{-1}})}\\
    \middlehline
    \chem{O_3}     & 1.6 & 1.0 & 1 & $10^{-2}$\\
    \chem{SO_2}    & 1.9 & 0.0 & 1 & $10^5$\\
    \chem{NO_2}    & 1.6 & 0.1 & 1 & $10^{-2}$\\
    \chem{H_2O_2}  & 1.4 & 1.0 & 1 & $10^5$\\
    \chem{HCHO}    & 1.3 & 0.0 & 6 & $10^3$\\
    \chem{NO}      & 1.3 & 0.0 & 2 & $10^{-3}$\\
    \chem{CH_3CHO} & 1.6 & 15.0 & \multicolumn{2}{l}{$0$}\\
    \bottomhline
  \end{tabular}
  %\belowtable{} % Table Footnotes
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{S.2 Dynamic viscosity of air and quasi-laminar resistance}
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity.pdf}
  \caption{Sutherland's law and divergence linear fits with one or two degrees of freedom. The fit $\mu(T) = m\cdot T$ is implemented in the OsloCTM3. Despite its strong divergence at lower and higher temperatures, the impact of this choice on the results is neglicible.}
\end{figure}
%
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight]{pictures/dynamic_viscosity_zo_Rb.pdf}
  \caption{Comparison of resulting $z_o$ and $R_b$ for $\mu(T) = m\cdot T$  (left panel) and Sutherland's law (right panel). The example displays the distributions for \chem{H_2O} after one day of model integration for both, rough and calm sea case.}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}[b]{1.\textwidth}
    \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_1.pdf}
    \caption{}
  \end{subfigure}
  \begin{subfigure}[b]{1.\textwidth}
    \includegraphics[height=0.45\textheight]{pictures/dynamic_viscosity_Rbi_2.pdf}
    \caption{}
  \end{subfigure}
  \caption{Resulting $R_b$ for different species. Shown are results of one day model integrations for both, Southerland's law and $\mu(T) = m\cdot T$. Shown are also the percentages above / below the given threashold.}
\end{figure}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{S.3 Stomatal conductance}
\begin{table*}[!htbp]
  \caption{Tabulated parameters for stomatal conductance computation in the EMEP scheme}

@@

  \begin{tabular}{lcccccccccc}

@@

    Code$^*$ & $D_\text{max}$ & $D_\text{min}$ & $D_\text{crit}$ & $R_\chem{SO}$ & $R_\chem{O_3}$ & $h$ & $d_\text{SGS}$ & $d_\text{EGS}$ & $\nabla d_\text{SGS}$ & $\nabla d_\text{EGS}$\\% & $\text{LAI}_\text{min}$ & $\text{LAI}_\text{max}$ & LS & LE \\
    &(\unit{kPA})&(\unit{kPa})&(\unit{kPa})&(\unit{s\,m^{-1}})&(\unit{s\,m^{-1}})&(\unit{m})&(\unit{day})&(\unit{day})&(\unit{days/^\circ lat})&(\unit{days/^\circ lat})\\%&(\unit{m^2\,m^{-2}})&(\unit{m^2\,m^{-2}})&(\unit{days})&(\unit{days}) \\

@@

    CF & 0.5 & 3 & 1000 & 0 & 200 & 20 & 0 & 366 & 0 & 0\\% & 5 & 5 & 1 & 1\\
    DF & 1 & 3.25 & 1000 & 0 & 200 & 20 & 100 & 307 & 1.5 & -2.0 \\%& 0 & 4 & 20 & 30\\
    NF & 1 & 3.2 & 1000 & 0 & 200 & 8 & 0 & 366 & 0 & 0\\% & 4 & 4 & 1 & 1\\
    BF & 2.2 & 4 & 1000 & 0 & 200 & 15 & 0 & 366 & 0 & 0 \\%& 4 & 4 & 1 & 1\\
    TC & 1.2 & 3.2 & 8 & 0 & 200 & 1 & 123 & 213 & 2.57 & 2.57\\% & 0 & 3.5 & 70 & 20\\
    MC & 1 & 2.5 & 1000 & 0 & 200 & 2 & 123 & 237 & 2.57 & 2.57\\% & 0 & 3 & 70 & 44\\
    RC & 0.31 & 2.7 & 10 & 0 & 200 & 1 & 130 & 250 & 0 & 0\\% & 0 & 4.2 & 35 & 65\\
    SNL & 1.3 & 3 & 1000 & 0 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 2 & 3 & 192 & 96\\
    GR & 1.3 & 3 & 1000 & 0 & 1000 & 0.3 & 0 & 366 & 0 & 0\\% & 2 & 3.5 & 140 & 135\\
    MS & 1.3 & 3.2 & 1000 & 0 & 200 & 2 & 0 & 366 & 0 & 0\\% & 2.5 & 2.5 & 1 & 1\\
    WE  & 1 & 0 & 1000 & 50 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
    TU & 1 & 0 & 1000 & 500 & 400 & 0.5 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
    DE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
    W  & 1 & 0 & 1000 & 1 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
    ICE & 1 & 0 & 1000 & 1000 & 2000 & 0 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\
    U & 1 & 0 & 1000 & 400 & 400 & 10 & 0 & 366 & 0 & 0\\% & 0 & 0 & 0 & 0\\

@@

\end{table*}

\clearpage
\subsection*{S.4 Latitude dependent vegetation height}
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight]{pictures/tree_height.pdf}
  \caption{Extension of the latitude dependent vegetation height from northern hemisphere mid latitudes only to a global describtion. Example heights are shown for coniferous, decideous, needleleaf, and broadleaf forests as well as Mediterranean shrub. As reference the original range of EMEP is added.}
\end{figure}

\subsection*{S.5 Temperature dependent greening season (python 2.7)}
{\scriptsize
\begin{verbatim}
#----------------------------------------------------------------------------------
import numpy as np
import pandas as pd
import xarray as xr
#----------------------------------------------------------------------------------
def start_growing_season_fixed(lat, **kwargs):
    '''
    Begin of growing season parameterization adapted from SMOKE-BEIS.
    '''
    bLeap = kwargs.pop('leap', False)
    if (bLeap):
       MAY31 = 152
       NOV1  = 306
       DEC31 = 366
    else:
       MAY31 = 151
       NOV1  = 305
       DEC31 = 365

@@

    return({'sgs':start_gs,'egs':end_gs})
#----------------------------------------------------------------------------------
def growing_season_stadyn(xr_temp):
    '''
    Preprocess the greening season for OsloCTM3.
    Setting using the 5deg-5days criteria and in case this fails
    or is out of the defined bounds,
    fall back to SMOKE-BEIS parameterization.
    '''
    # Fetch leap year from input data
    bLeap = False
    if xr_temp.time.size==366:
       bLeap = True
    # Use 5deg-5days criteria   
    if (xr_temp.lat > 45 and xr_temp.lat < 85):
        gs = growing_season(xr_temp)
        sgs = (gs['sgs'][0],)
        egs = (gs['egs'][0],)
        if ( not gs['sgs'][1] or not gs['egs'][1] or sgs[0]>=egs[0]):
            # Check for failing 5deg-5days criteria => Fall back to fixed
            sgs = start_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
            egs = end_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
    elif (xr_temp.lat < -35 and xr_temp.lat >= -65):
        gs = growing_season(xr_temp, sh=True)
        sgs = (gs['sgs'][0],1)
        egs = (xr_temp.size, gs['egs'][0])
        # Check for failing 5deg-5days criteria => Fall back to fixed
        if (not gs['sgs'][1] or not gs['egs'][1] or sgs[0]>=egs[0]):
            sgs = start_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
            egs = end_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
    # Use SMOKE-BEIS parameterization
    else:
        sgs = start_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
        egs = end_growing_season_fixed(xr_temp.lat.data, leap=bLeap)
    # Generate GDAY and GLEN fields
    # in case no growing season has been allocated
    if (sgs[0]==egs[0]==0):
        gday = np.repeat(0,len(xr_temp.time))
        glen = 0
    # in normal cases
    else:
        # handle northern hemisphere
        if len(sgs)==1:
            gday = np.concatenate((np.repeat(0,int(sgs[0])-1),
                                   np.arange(1,int(egs[0])-int(sgs[0])+1),
                                   np.repeat(0,len(xr_temp.time)-
                                   int(egs[0])+1)))
            glen = int(egs[0])-int(sgs[0])+1
        # handle southern hemisphere
        else:
            gday = np.concatenate((np.arange(1,int(egs[1])+1)+len(xr_temp.time)
                                   -int(sgs[0])+1,
                                   np.repeat(0,int(sgs[0])-int(egs[1])),
                                   np.arange(1,len(xr_temp.time)-int(sgs[0])+1)))
            glen = len(xr_temp.time)-(int(sgs[0])-int(egs[1]))+1
    # This should not happen, but in case it does, print info.
    if not(len(gday)==len(xr_temp.time)):
        print(xr_temp, sgs, egs)
    # Output
    data_gday = xr.DataArray(gday.astype(int),[('time', xr_temp.time.data)])
    return(data_gday, (int)(glen))
#----------------------------------------------------------------------------------
\end{verbatim}
}
\clearpage

\subsection*{S.6 De-accumulation of photosynthetic active radiation (PAR) from OpenIFS}
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.65cm}]{pictures/PAR_as_is.png}
  \caption{Example output from OpenIFS for January 2nd 2005. PAR is accumulated over the period of one day.}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.65cm}]{pictures/PAR_partly_deaccumulated.png}
  \caption{Partly de-accumulated output from OpenIFS. $\text{PPFD}(t_i) = \text{PAR}(t_{i+1})-\text{PAR}(t_i)$.}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight, clip, trim={0.75cm 0.75cm 0.75cm 0.65cm}]{pictures/PAR_deaccumulating_00h.png}
  \caption{De-accumulation of hour 00 in the partly de-accumulated fields. $\text{PPFD}(t=0) = \text{PAR}(t=0) - \left[\text{PAR}(t=21\,\unit{h}-1\,\unit{day})-\text{PAR}(=12\,\unit{h}-1\,\unit{day})\right]$.}
\end{figure}
%\clearpage

\subsection*{S.7 Normalized zonal average ozone dry deposition velocity}
\begin{figure}[!htbp]
  \centering
  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-norm_total_ozone_drydepvelo}
  \caption{Normalized zonal average ozone dry deposition velocity.}
\end{figure}
